---
title: "Call cell types"
author: "Zhaoheng Li, Wei Sun"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: journal
    highlight: tango
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: false
    number_sections: false
  df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R library and data input

```{r load_libraries, warning = FALSE, message = FALSE}
library(Seurat)
library(Matrix)
library(dplyr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(tibble)
library(cowplot)  
library(patchwork)
library(stringr)
library(pheatmap)
library(reshape2)
library(readxl)
library(patchwork)

theme_set(theme_classic())

base_dir = "/fh/working/sun_w/zli/MSK_DEC24"
data_dir = "/fh/working/sun_w/zli/MSK_DEC24/data2/processed_samples"
output_dir = "/fh/working/sun_w/wsun/MSK_KO_village_cell_typing/output"

setwd(base_dir)
```

## Read in Seurat objects

```{r}
integrated = readRDS(file.path(output_dir,"integrated-rpca_BDF-ref.RDS"))
dim(integrated)
```


## Update meta data information

Focus on clusters at resolution 0.5

```{r}
meta = integrated[[]]
dim(meta)
meta[1:2,]

meta <- meta %>% select(-mt_counts, -mt_pct, 
                        -n_genes_by_counts, 
                        -log1p_n_genes_by_counts, 
                        -total_counts, 
                        -log1p_total_counts, 
                        -pct_counts_in_top_50_genes, 
                        -pct_counts_in_top_100_genes, 
                        -pct_counts_in_top_200_genes, 
                        -pct_counts_in_top_500_genes, 
                        -log1p_total_counts_mt, 
                        -seurat_clusters,
                        -RNA_snn_res.1, 
                        -RNA_snn_res.0.8, 
                        -RNA_snn_res.0.3)

dim(meta)
meta[1:2,]

meta$barcode = rownames(meta)

summary(meta$pct_counts_mt)
summary(meta$doublet_scores)
```

## Check the sizes of clusters
```{r}
df_cluster <- data.frame(table(integrated$RNA_snn_res.0.5))
names(df_cluster)[1] = "cluster"
df_cluster$cluster = as.numeric(as.character(df_cluster$cluster))
df_cluster = df_cluster[order(df_cluster$cluster),]
rownames(df_cluster) = NULL
dim(df_cluster)
df_cluster

cat(df_cluster$Freq, sep=",")
```


## Check a few group of DEs

```{r}
find_markers <- function(seurat_obj, cluster1, cluster2) {
  # Step 1: Wilcoxon test
  wilcox_res <- FindMarkers(
    seurat_obj,
    ident.1 = cluster1,
    ident.2 = cluster2,
    test.use = "wilcox",
    logfc.threshold = 0,
    min.pct = 0
  )
  
  # Step 2: Filter by |log2FC| > 1
  filtered_genes <- wilcox_res[abs(wilcox_res$avg_log2FC) > 1, ]
  
  # Step 3: Select top 1000 genes by adjusted p-value
  ng = min(1000, nrow(filtered_genes))
  top_genes <- rownames(head(filtered_genes[order(filtered_genes$p_val_adj), ], ng))
  
  # Step 4: ROC test on top genes
  roc_res <- FindMarkers(
    seurat_obj,
    ident.1 = cluster1,
    ident.2 = cluster2,
    test.use = "roc",
    features = top_genes
  )
  
  # Step 5: Filter genes with AUC > 0.8
  roc_filtered <- roc_res[roc_res$myAUC > 0.7 | roc_res$myAUC < 0.3, ]
  
  # Step 6: Select top 10 with largest pct.1 - pct.2
  top10 = min(10, nrow(roc_filtered))
  roc_filtered$delta <- roc_filtered$pct.1 - roc_filtered$pct.2
  top_cluster1 <- head(roc_filtered[order(-roc_filtered$delta), ], top10)
  top_cluster2 <- head(roc_filtered[order(roc_filtered$delta), ], top10)
  
  # Return list of results
  return(list(
    wilcox_all = wilcox_res,
    roc_all = roc_res,
    roc_filtered = roc_filtered,
    top_cluster1_markers = top_cluster1,
    top_cluster2_markers = top_cluster2
  ))
}
```

### Distinguish cluster 4 (ductal) and 12 (liver).

```{r}
table(integrated$RNA_snn_res.0.5)
Idents(integrated) <- "RNA_snn_res.0.5"

sc_4_vs_12 <- find_markers(integrated, 4, 12)

sc_4_vs_12$top_cluster1_markers
sc_4_vs_12$top_cluster2_markers
```

### Distinguish cluster 3 (sc-EC) and 8 (sc-alpha).

```{r}
table(integrated$RNA_snn_res.0.5)
Idents(integrated) <- "RNA_snn_res.0.5"

sc_3_vs_8 <- find_markers(integrated, 3, 8)

sc_3_vs_8$top_cluster1_markers
sc_3_vs_8$top_cluster2_markers
```

### Makers of each cluster

```{r}

for(c1 in 0:20){
  markers <- FindMarkers(integrated, 
                         ident.1 = c1, 
                         ident.2 = NULL, 
                         only.pos = TRUE, 
                         logfc.threshold = 0.5, 
                         min.pct = 0.5)
  print(c1)
  print(head(markers))
}

```

## Check marker genes

Mannual examination of the markers listed above plus additional markers from literature. 

ESC -> DE (Definitive endoderm) -> Primitive Gut Tube -> PFG (Posterior Foregut) 

PFG -> Pancreatic progenitors -> beta/alpha/duct cells


```{r}
celltype_markers = list(
  ESC                 = c("POU5F1", "DPPA4", "NANOG", "ESRG"),
  DE                  = c("SOX17", "CXCR4", "EOMES", "CYP26A1", "HAS2"),
  Primitive_Gut_tube  = c("APELA", "SMAD6", "PCAT14", "SEMA3E", "ADAMTS12"),
  PFG                 = c("ONECUT1", "CDH6", "ZFPM2", "CXXC4"),
  PP                  = c("SLC4A4", "SPINK1", "DLK1", "PDX1", "SOX9"),
  Mesenchymal_muscle  = c("ANKRD1", "TAGLN", "ANXA3", "MYL9", "TNFRSF12A"),
  Endocrine_precursor = c("NEUROG3", "INSM1", "FEV"),
  SC_beta             = c("ACVR1C", "CALB2", "ISL1", "MAFB", "ABCC8", "INS"),
  SC_alpha            = c("GCG", "ARX", "ABTB3", "ISL1"),
  SC_delta            = c("SST", "HHEX"),
  SC_enterchromaffin  = c("TPH1", "LMX1A", "FEV"),
  Liver               = c("AFP", "RBP4", "APOA2"),
  Ductal              = c("THRB", "KLF5", "ABCG2", "GSTA1", "LGALS3"),
  Stromal             = c("COL1A1", "COL3A1", "TFAP2A", "ANXA1", "SPARC"),
  Endothelial         = c("PLVAP", "PECAM1", "FLT1", "GNG11", "CDH5")
)

celltype_markers = lapply(celltype_markers, function(markers) {
  intersect(markers, rownames(integrated))
})
celltype_markers

for (celltype in names(celltype_markers)) {
  genes <- celltype_markers[[celltype]]
  
  # Filter to genes actually present in the data
  genes <- genes[genes %in% rownames(integrated)]
  
  # Further filter out genes expressed in ≤ 0.1% of cells
  genes <- genes[sapply(genes, function(gene) {
    expr <- FetchData(integrated, vars = gene)[[1]]
    mean(expr > 1) > 0.001
  })]
  
  n_genes <- length(genes)

  if (n_genes == 0) next
  
  cat("\n\n###", celltype, "\n\n")
  
  # Dynamically adjust panel layout
  n_cols <- 3
  n_rows <- ceiling(n_genes / n_cols)
  
  # Generate individual FeaturePlots
  plots <- lapply(genes, function(gene) {
    FeaturePlot(integrated, features = gene, reduction = "rpca.umap") + 
      ggtitle(gene)
  })
  
  # Combine into a patchwork layout and print
  plot_grid <- wrap_plots(plots, ncol = n_cols)
  
  pdf_file <- file.path(output_dir, "cell_type_markers", 
                        paste0("markers_", celltype, ".pdf"))
  pdf(pdf_file, width = 4 * n_cols, height = 4 * n_rows)
  print(plot_grid)
  dev.off()
}

```

## Check marker genes for cluster 10

```{r}
genes = c("SCG2", "GC", "C1QL1", "STMN2", "ONECUT3", "SEZ6L")
n_cols <- 3
n_rows <- ceiling(n_genes / n_cols)

plots <- lapply(genes, function(gene) {
  FeaturePlot(integrated, features = gene, reduction = "rpca.umap") + 
    ggtitle(gene)
})

plot_grid <- wrap_plots(plots, ncol = n_cols)

pdf_file <- file.path(output_dir, "cell_type_markers", 
                      paste0("markers_cluster10.pdf"))
pdf(pdf_file, width = 4 * n_cols, height = 4 * n_rows)
print(plot_grid)
dev.off()
```
  
## Draw a heatmap of marker genes across clusters

```{r}
all_markers <- unique(unlist(celltype_markers))

# Get cluster IDs
meta$cluster <- meta$RNA_snn_res.0.5
Idents(integrated) <- meta$cluster

# Compute average expression
avg_expr <- AggregateExpression(integrated, features = all_markers, 
                              group.by = "RNA_snn_res.0.5", 
                              return.seurat = FALSE)$RNA
dim(avg_expr)
avg_expr[1:2,]

summary(colSums(avg_expr))

lib_size <- colSums(avg_expr)
norm_expr <- sweep(avg_expr, 2, lib_size, FUN = "/") * median(colSums(avg_expr))

log_expr <- log2(norm_expr + 1)
z_expr <- t(scale(t(log_expr)))

# Compute percent of cells expressing each gene in each cluster
pct_expr <- sapply(all_markers, function(gene) {
  tapply(FetchData(integrated, vars = gene)[,1] > 0, meta$cluster, mean)
}) %>% t()

# Reshape for plotting
df_expr <- z_expr %>%
  as.data.frame() %>%
  rownames_to_column("gene") %>%
  pivot_longer(-gene, names_to = "cluster", values_to = "avg_expr")

dim(df_expr)
df_expr[1:5,]
df_expr$cluster = gsub("g", "", df_expr$cluster)

df_pct <- as.data.frame(pct_expr) %>%
  rownames_to_column("gene") %>%
  pivot_longer(-gene, names_to = "cluster", values_to = "pct_expr")

dim(df_pct)
df_pct[1:5,]

df_plot <- left_join(df_expr, df_pct, by = c("gene", "cluster"))

dim(df_plot)
df_plot[1:5,]

pf_pct_wider = df_pct %>%
  tidyr::pivot_wider(names_from = cluster, values_from = pct_expr)

gene_order <- pf_pct_wider %>%
  column_to_rownames("gene") %>%
  dist() %>%
  hclust() %>%
  as.dendrogram() %>%
  order.dendrogram()

# Reorder gene factor levels
lvs = pf_pct_wider$gene[gene_order]

df_plot$gene <- factor(df_plot$gene, levels = lvs)

df_expr_wider = df_expr %>%
  pivot_wider(names_from = cluster, values_from = avg_expr, values_fill = 0) %>% 
  column_to_rownames("gene")

dim(df_expr_wider)
df_expr_wider[1:2,]

cluster_order <- df_expr_wider  %>%
  t() %>%
  dist() %>%
  hclust() %>%
  as.dendrogram() %>%
  order.dendrogram()

# Apply the new cluster order
ordered_clusters <- colnames(df_expr_wider)[cluster_order]
df_plot$cluster <- factor(df_plot$cluster, levels = ordered_clusters)


pdf(file.path(output_dir, "figures/cluster_vs_markers_heatmap.pdf"), 
    width = 6, height = 9)
ggplot(df_plot, aes(x = cluster, y = gene)) +
  geom_point(aes(size = pct_expr, color = avg_expr)) + 
  scale_size(range = c(1, 4), name = "Pct Expr") + 
  scale_color_gradient2(low = "steelblue", mid = "white", high = "red", 
                        midpoint = 0,  # Assumes z-scored avg_expr centered at 0
                        name = "Avg Expr") +
  theme_minimal() +
  labs(x = "Cluster", y = "Marker Gene",
       size = "Pct Expr", color = "Avg Expr") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
dev.off()

```

## Function to add subcluster information

```{r}
update_sub_cluster <- function(sub_obj, cluster_id, possible_cell_types, 
                               celltype_markers){

  all_markers <- NULL
  for(ct1 in possible_cell_types){
    all_markers = c(all_markers, celltype_markers[[ct1]])
  }
  
  ## remove genes that are markers for more than 1 cell type
  all_markers <- names(table(all_markers)[table(all_markers) == 1])
  all_markers <- all_markers[all_markers %in% rownames(sub_obj)]

  avg_exp <- AverageExpression(sub_obj, features = all_markers, 
                               return.seurat = FALSE)$RNA
  dim(avg_exp)
  
  gene_to_celltype = rep(NA, nrow(avg_exp))
  for(ct1 in possible_cell_types){
    ww1 = which(rownames(avg_exp) %in% celltype_markers[[ct1]])
    gene_to_celltype[ww1] = ct1
  }
  
  avg_exp = data.frame(avg_exp)
  avg_exp$marker_group <- gene_to_celltype
  
  avg_long <- avg_exp %>%
    tibble::rownames_to_column("gene") %>%
    pivot_longer(-c(gene, marker_group), names_to = "subcluster", values_to = "avg_expr")
  
  pdf_file <- file.path(output_dir, 
                        sprintf("recluster/cluster%s/marker_genes.pdf", cluster_id))
  pdf(pdf_file, width = 6, height = 4)
  g1 = ggplot(avg_long, aes(x = marker_group, y = avg_expr, fill = subcluster)) +
    geom_boxplot() + 
    theme_minimal()
  print(g1)
  dev.off()
}
```

## Re-clustering cluster 2

```{r}
meta$subcluster <- NA

Idents(integrated) <- "RNA_snn_res.0.5"

sub_obj <- subset(integrated, idents = c("2"))

sub_obj <- sub_obj %>% 
  NormalizeData() %>% 
  FindVariableFeatures() %>% 
  ScaleData() %>% 
  RunPCA(npcs = 30)

sub_obj <- sub_obj %>% 
  FindNeighbors(dims = 1:20) %>% 
  FindClusters(resolution = 0.1)

sub_obj <- RunUMAP(sub_obj, dims = 1:20)

dir1 = file.path(output_dir, "recluster/cluster2")
if (!dir.exists(dir1)) { dir.create(dir1) }

pdf(file.path(output_dir, "recluster/cluster2", "umap_by_clusters.pdf"), 
    width = 6, height = 5)
DimPlot(sub_obj, group.by = "seurat_clusters", label = TRUE) + 
  ggtitle("Subclusters of Cluster 2")
dev.off()

cluster_id = "2"
possible_cell_types = c("SC_enterchromaffin", "Endocrine_precursor")

Idents(sub_obj) <- "seurat_clusters"
table(Idents(sub_obj))

new_labels <- paste(cluster_id, Idents(sub_obj), sep = "_")
names(new_labels) <- names(Idents(sub_obj)) 
new_labels[1:5]

mat1 = match(names(new_labels), meta$barcode)
stopifnot(sum(is.na(mat1)) == 0)
meta$subcluster[mat1] <- new_labels
table(meta$subcluster)

update_sub_cluster(sub_obj, cluster_id, possible_cell_types, celltype_markers)
```

## Re-clustering cluster 16

```{r}
sub_obj <- subset(integrated, idents = c("16"))

sub_obj <- sub_obj %>% 
  NormalizeData() %>% 
  FindVariableFeatures() %>% 
  ScaleData() %>% 
  RunPCA(npcs = 30)

sub_obj <- sub_obj %>% 
  FindNeighbors(dims = 1:20) %>% 
  FindClusters(resolution = 0.1)

sub_obj <- RunUMAP(sub_obj, dims = 1:20)

dir1 = file.path(output_dir, "recluster/cluster16")
if (!dir.exists(dir1)) { dir.create(dir1) }

pdf(file.path(output_dir, "recluster/cluster16", "umap_by_clusters.pdf"), 
    width = 6, height = 5)
DimPlot(sub_obj, group.by = "seurat_clusters", label = TRUE) + 
  ggtitle("Subclusters of Cluster 16")
dev.off()

cluster_id = "16"
possible_cell_types = c("DE", "Primitive_Gut_tube")

Idents(sub_obj) <- "seurat_clusters"
table(Idents(sub_obj))

new_labels <- paste(cluster_id, Idents(sub_obj), sep = "_")
names(new_labels) <- names(Idents(sub_obj)) 
new_labels[1:5]

mat1 = match(names(new_labels), meta$barcode)
stopifnot(sum(is.na(mat1)) == 0)
meta$subcluster[mat1] <- new_labels
table(meta$subcluster)

update_sub_cluster(sub_obj, cluster_id, possible_cell_types, celltype_markers)
```


## Re-clustering cluster 5

```{r}
sub_obj <- subset(integrated, idents = c("5"))

sub_obj <- sub_obj %>% 
  NormalizeData() %>% 
  FindVariableFeatures() %>% 
  ScaleData() %>% 
  RunPCA(npcs = 30)

sub_obj <- sub_obj %>% 
  FindNeighbors(dims = 1:20) %>% 
  FindClusters(resolution = 0.1)

sub_obj <- RunUMAP(sub_obj, dims = 1:20)

dir1 = file.path(output_dir, "recluster/cluster5")
if (!dir.exists(dir1)) { dir.create(dir1) }

pdf(file.path(output_dir, "recluster/cluster5", "umap_by_clusters.pdf"), 
    width = 6, height = 5)
DimPlot(sub_obj, group.by = "seurat_clusters", label = TRUE) + 
  ggtitle("Subclusters of Cluster 5")
dev.off()

cluster_id = "5"
possible_cell_types = c("PP", "Ductal", "Liver")

Idents(sub_obj) <- "seurat_clusters"
table(Idents(sub_obj))

new_labels <- paste(cluster_id, Idents(sub_obj), sep = "_")
names(new_labels) <- names(Idents(sub_obj)) 
new_labels[1:5]

mat1 = match(names(new_labels), meta$barcode)
stopifnot(sum(is.na(mat1)) == 0)
meta$subcluster[mat1] <- new_labels
table(meta$subcluster)

update_sub_cluster(sub_obj, cluster_id, possible_cell_types, celltype_markers)
```

## Re-clustering cluster 10

```{r}
sub_obj <- subset(integrated, idents = c("10"))

sub_obj <- sub_obj %>% 
  NormalizeData() %>% 
  FindVariableFeatures() %>% 
  ScaleData() %>% 
  RunPCA(npcs = 30)

sub_obj <- sub_obj %>% 
  FindNeighbors(dims = 1:20) %>% 
  FindClusters(resolution = 0.1)

sub_obj <- RunUMAP(sub_obj, dims = 1:20)

dir1 = file.path(output_dir, "recluster/cluster10")
if (!dir.exists(dir1)) { dir.create(dir1) }

pdf(file.path(output_dir, "recluster/cluster10", "umap_by_clusters.pdf"), 
    width = 6, height = 5)
DimPlot(sub_obj, group.by = "seurat_clusters", label = TRUE) + 
  ggtitle("Subclusters of Cluster 10")
dev.off()


cluster_id = "10"
possible_cell_types = c("SC_alpha", "Endocrine_precursor", 
                        "SC_enterchromaffin")

Idents(sub_obj) <- "seurat_clusters"
table(Idents(sub_obj))

new_labels <- paste(cluster_id, Idents(sub_obj), sep = "_")
names(new_labels) <- names(Idents(sub_obj)) 
new_labels[1:5]

mat1 = match(names(new_labels), meta$barcode)
stopifnot(sum(is.na(mat1)) == 0)
meta$subcluster[mat1] <- new_labels
table(meta$subcluster)

update_sub_cluster(sub_obj, cluster_id, possible_cell_types, celltype_markers)
```


## Re-clustering cluster 3

Just to get a reference for the reclustering of cluster 10.

```{r}
sub_obj <- subset(integrated, idents = c("3"))

sub_obj <- sub_obj %>% 
  NormalizeData() %>% 
  FindVariableFeatures() %>% 
  ScaleData() %>% 
  RunPCA(npcs = 30)

sub_obj <- sub_obj %>% 
  FindNeighbors(dims = 1:20) %>% 
  FindClusters(resolution = 0.2)

sub_obj <- RunUMAP(sub_obj, dims = 1:20)

dir1 = file.path(output_dir, "recluster/cluster3")
if (!dir.exists(dir1)) { dir.create(dir1) }

pdf(file.path(output_dir, "recluster/cluster3", "umap_by_clusters.pdf"), 
    width = 6, height = 5)
DimPlot(sub_obj, group.by = "seurat_clusters", label = TRUE) + 
  ggtitle("Subclusters of Cluster 3")
dev.off()


cluster_id = "3"
possible_cell_types = c("Endocrine_precursor", "SC_alpha", "SC_enterchromaffin")

Idents(sub_obj) <- "seurat_clusters"
table(Idents(sub_obj))

update_sub_cluster(sub_obj, cluster_id, possible_cell_types, celltype_markers)
```

## Re-clustering cluster 8

Just to get a reference for the reclustering of cluster 10.

```{r}
sub_obj <- subset(integrated, idents = c("8"))

sub_obj <- sub_obj %>% 
  NormalizeData() %>% 
  FindVariableFeatures() %>% 
  ScaleData() %>% 
  RunPCA(npcs = 30)

sub_obj <- sub_obj %>% 
  FindNeighbors(dims = 1:20) %>% 
  FindClusters(resolution = 0.2)

sub_obj <- RunUMAP(sub_obj, dims = 1:20)

dir1 = file.path(output_dir, "recluster/cluster8")
if (!dir.exists(dir1)) { dir.create(dir1) }

pdf(file.path(output_dir, "recluster/cluster8", "umap_by_clusters.pdf"), 
    width = 6, height = 5)
DimPlot(sub_obj, group.by = "seurat_clusters", label = TRUE) + 
  ggtitle("Subclusters of Cluster 8")
dev.off()


cluster_id = "8"
possible_cell_types = c("Endocrine_precursor", "SC_alpha", "SC_enterchromaffin")

Idents(sub_obj) <- "seurat_clusters"
table(Idents(sub_obj))

update_sub_cluster(sub_obj, cluster_id, possible_cell_types, celltype_markers)
```

## Re-clustering cluster 11

```{r}
sub_obj <- subset(integrated, idents = c("11"))

sub_obj <- sub_obj %>% 
  NormalizeData() %>% 
  FindVariableFeatures() %>% 
  ScaleData() %>% 
  RunPCA(npcs = 30)

sub_obj <- sub_obj %>% 
  FindNeighbors(dims = 1:20) %>% 
  FindClusters(resolution = 0.2)

sub_obj <- RunUMAP(sub_obj, dims = 1:20)

dir1 = file.path(output_dir, "recluster/cluster11")
if (!dir.exists(dir1)) { dir.create(dir1) }

pdf(file.path(output_dir, "recluster/cluster11", "umap_by_clusters.pdf"), 
    width = 6, height = 5)
DimPlot(sub_obj, group.by = "seurat_clusters", label = TRUE) + 
  ggtitle("Subclusters of Cluster 11")
dev.off()


cluster_id = "11"
possible_cell_types = c("Endocrine_precursor", "SC_alpha", "SC_enterchromaffin")

Idents(sub_obj) <- "seurat_clusters"
table(Idents(sub_obj))

new_labels <- paste(cluster_id, Idents(sub_obj), sep = "_")
names(new_labels) <- names(Idents(sub_obj)) 
new_labels[1:5]

mat1 = match(names(new_labels), meta$barcode)
stopifnot(sum(is.na(mat1)) == 0)
meta$subcluster[mat1] <- new_labels
table(meta$subcluster)

update_sub_cluster(sub_obj, cluster_id, possible_cell_types, celltype_markers)
```

## Re-clustering cluster 15

```{r}
Idents(integrated) <- "RNA_snn_res.0.5"

sub_obj <- subset(integrated, idents = c("15"))

sub_obj <- sub_obj %>% 
  NormalizeData() %>% 
  FindVariableFeatures() %>% 
  ScaleData() %>% 
  RunPCA(npcs = 30)

sub_obj <- sub_obj %>% 
  FindNeighbors(dims = 1:20) %>% 
  FindClusters(resolution = 0.1)

sub_obj <- RunUMAP(sub_obj, dims = 1:20)

dir1 = file.path(output_dir, "recluster/cluster15")
if (!dir.exists(dir1)) { dir.create(dir1) }

pdf(file.path(output_dir, "recluster/cluster15", "umap_by_clusters.pdf"), 
    width = 6, height = 5)
DimPlot(sub_obj, group.by = "seurat_clusters", label = TRUE) + 
  ggtitle("Subclusters of Cluster 15")
dev.off()

cluster_id = "15"
possible_cell_types = c("Ductal", "PP")

Idents(sub_obj) <- "seurat_clusters"
table(Idents(sub_obj))

new_labels <- paste(cluster_id, Idents(sub_obj), sep = "_")
names(new_labels) <- names(Idents(sub_obj)) 
new_labels[1:5]

mat1 = match(names(new_labels), meta$barcode)
stopifnot(sum(is.na(mat1)) == 0)
meta$subcluster[mat1] <- new_labels
table(meta$subcluster)

update_sub_cluster(sub_obj, cluster_id, possible_cell_types, celltype_markers)
```

### Update meta data

```{r}
dim(meta)
meta[1:2,]

table(meta$feature.4 == meta$feature.10, useNA = 'ifany')


meta$celltype = NA

meta <- meta %>%
  mutate(
    celltype = case_when(
      RNA_snn_res.0.5 == 0  ~ "ESC",
      RNA_snn_res.0.5 == 1  ~ "DE",
      RNA_snn_res.0.5 == 3  ~ "SC-EC",
      RNA_snn_res.0.5 == 4  ~ "Ductal",
      RNA_snn_res.0.5 == 6  ~ "SC-beta",
      RNA_snn_res.0.5 == 7  ~ "PP",
      RNA_snn_res.0.5 == 8  ~ "SC-alpha",
      RNA_snn_res.0.5 == 9  ~ "PFG",
      RNA_snn_res.0.5 == 11 ~ "EnP",
      RNA_snn_res.0.5 == 12 ~ "Liver",
      RNA_snn_res.0.5 == 13 ~ "ESC_DE",
      RNA_snn_res.0.5 == 14 ~ "Stromal",
      RNA_snn_res.0.5 == 15 ~ "Ductal",
      RNA_snn_res.0.5 == 17 ~ "SC-delta",
      RNA_snn_res.0.5 == 18 ~ "DE",
      RNA_snn_res.0.5 == 19 ~ "Mesenchymal_Stromal",
      RNA_snn_res.0.5 == 20 ~ "Endothelial",
      subcluster %in% c("2_0", "2_1", "2_2") ~ "EnP",
      subcluster == "2_3"                    ~ "DE",
      subcluster %in% c("5_0", "5_3") ~ "PP",
      subcluster %in% c("5_2") ~ "Ductal",
      subcluster %in% c("5_1") ~ "Liver",
      subcluster %in% c("10_3") ~ "SC-EC",
      subcluster %in% c("16_0", "16_1", "16_2") ~ "PGT",
      TRUE ~ NA_character_
    )
  )

table(meta$RNA_snn_res.0.5, meta$celltype, useNA = "ifany")
table(meta$subcluster, meta$celltype, useNA = "ifany")

write.csv(meta, 
          file.path(output_dir, "MSK_KO_village_meta_data.csv"), 
          row.names = FALSE)

```

# Session information
```{r}
gc()

sessionInfo()
```

