---
title: "Integrate external WT and KO village"
author: "Zhaoheng Li, Wei Sun"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: journal
    highlight: tango
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: false
    number_sections: false
  df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R library and data input

```{r load_libraries, warning = FALSE, message = FALSE}
library(Seurat)
library(Matrix)
library(dplyr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(tibble)
library(cowplot)  
library(patchwork)
library(stringr)
library(pheatmap)
library(reshape2)
library(readxl)

theme_set(theme_classic())

base_dir = "/fh/working/sun_w/zli/MSK_DEC24"
data_dir = "/fh/working/sun_w/zli/MSK_DEC24/data2/processed_samples"
output_dir = "/fh/working/sun_w/wsun/MSK_KO_village_cell_typing/output"

setwd(base_dir)
```

## Read in Seurat objects

Use samples external wildtype samples B, D, and F (entries 2, 4, and 6 in ```seurat_list```) as reference for integration because in later time points, these 3 samples use 3D culture that are similar to the KO village. 

```{r}
seurat_list = readRDS(file.path(output_dir,"seurat_list.RDS"))
names(seurat_list)
length(seurat_list)

# merge them in this order so that the ref is the first object

ref =  merge(seurat_list[[2]], y = seurat_list[c(4,6)])

merged_all = merge(ref, y = seurat_list[setdiff(1:13,c(2,4,6))])

merged_all $project_id = with(merged_all @meta.data, ifelse(
  sample_name %in% c("B", "D", "F"),
  "ref",
  sample_name
))

merged_all$project_id = as.factor(merged_all$project_id)
merged_all[['RNA']] = JoinLayers(merged_all[['RNA']],layer = 'counts')
merged_all[["RNA"]] = split(merged_all[["RNA"]], f = merged_all$project_id)
merged_all

merged_all = NormalizeData(merged_all)
merged_all = FindVariableFeatures(merged_all, nfeatures = 2000)
merged_all = ScaleData(merged_all)

```


## Load gene annoation

```{r load gene annoation}
gene_anno_file = "gencode_v44_primary_assembly_info.tsv"

assembly_info = read.delim(
  file      = file.path(base_dir, "gene_annotation/v44", gene_anno_file),
  header    = TRUE,
  sep       = "\t",
  stringsAsFactors = FALSE
)

dim(assembly_info)
assembly_info[1:2,]

sum(is.na(assembly_info$hgnc_symbol))
sum(assembly_info$hgnc_symbol=='', na.rm = T)

assembly_info = assembly_info %>%
    mutate(hgnc_symbol = if_else(is.na(hgnc_symbol) | hgnc_symbol == "", 
                               ensembl_gene_id, 
                               hgnc_symbol))
sum(is.na(assembly_info$hgnc_symbol))
sum(assembly_info$hgnc_symbol=='', na.rm= T)
sum(assembly_info$hgnc_symbol == assembly_info$ensembl_gene_id)
```

## match ensembl ID to gene symbols

```{r}
symbol_map = setNames(
  assembly_info$hgnc_symbol,
  assembly_info$ensembl_gene_id
)

feat_meta = merged_all@assays$RNA@meta.data
rownames(feat_meta) = rownames(merged_all)

dim(feat_meta)
names(feat_meta)
feat_meta[1:2,1:4]

feat_meta$hgnc_symbol = symbol_map[rownames(feat_meta)]
merged_all@assays$RNA@meta.data = feat_meta
merged_all@assays$RNA@meta.data$ensembl_id = rownames(merged_all@assays$RNA@meta.data)
sum(is.na(feat_meta$hgnc_symbol)) 

# Get the hgnc_symbol vector
ids = as.character(merged_all@assays$RNA@meta.data$hgnc_symbol)

# Find duplicated entries (both forward and backward)
duplicated_ids = ids[duplicated(ids) | duplicated(ids, fromLast = TRUE)]
duplicated_ids_unique = unique(duplicated_ids)
duplicated_ids_unique

# set row names
rownames(merged_all[["RNA"]]) = make.unique(merged_all@assays$RNA@meta.data$hgnc_symbol)
duplicated_ids_unique %in% VariableFeatures(merged_all)
```

## PCA analysi

Here are the interpretation of PCs based on top positive/negative genes by ChatGPT. 


| **PC** | **Biological Axis**                      | **Positive Pole**             | **Negative Pole**                |
|--------|------------------------------------------|-------------------------------|----------------------------------|
| PC1    | Endocrine vs Pluripotent                 | INS, CHGA, SCGN               | POU5F1, LIN28A                   |
| PC2    | Mesenchymal/Epithelial vs Pluripotent    | FN1, CDH6                     | CRIPTO, DNMT3B                   |
| PC3    | Definitive Endoderm vs Pluripotent       | SOX17, CER1                   | MYC, MT1 genes                   |
| PC4    | Neuronal vs Hepatic                      | DLK1, AUTS2                   | APOE, AFP                        |
| PC5    | Neuronal Signaling vs Gut/Endocrine      | PLXDC2, GRIA4                 | NEUROG3, GHRL                    |

```{r}
merged_all = RunPCA(merged_all)
merged_all = RunUMAP(merged_all , dims = 1:30)

merged_all@assays$RNA

merged_all$time_point = as.factor(merged_all$time_point)
summary(merged_all$time_point)

new_levels = c(
  "D-1",  "D3",  "D5",  "D7",  "D9",  "D10", "D11",
  "D12", "D13", "D14", "D15", "D16", "D18", "D21",
  "D25", "D26", "D35"
)
merged_all$time_point = factor(merged_all$time_point,
                                levels = new_levels)

```


## merge culture_condition categories

```{r}
cc = as.character(merged_all@meta.data$culture_condition)
cc = str_trim(cc)                   # drop stray whitespace
cc = str_replace_all(cc, "–", "-")  # normalize any en–dash/chars

cc2 = as.factor(ifelse(
  is.na(cc), 
  "unannotated",
  ifelse(
    str_detect(cc, regex("\\b3D\\b", ignore_case = TRUE)), "3D",
    ifelse(
      str_detect(cc, regex("\\b2D\\b", ignore_case = TRUE)), "2D",
      cc
    )
  )
))

merged_all = AddMetaData(
  object    = merged_all,
  metadata  = cc2,
  col.name  = "culture_condition2"
)

merged_all$source = ifelse(
  merged_all$sample_name %in% c("A","B","C","D","E","F"),
  "external_WT",
  "KO_village"
)
```

## Integration

```{r}
lv_all = levels(merged_all$project_id)
merged_all$project_id <- factor(merged_all$project_id, 
                                levels = c("ref",  setdiff(lv_all, "ref")))
levels(merged_all$project_id)
merged_all

integrated = IntegrateLayers(object = merged_all, method = RPCAIntegration, 
                             orig.reduction = "pca", 
                             new.reduction = "integrated.rpca", 
                             verbose = TRUE, reference = c(1))

# flattens the data structure for standard Seurat operations
integrated[["RNA"]] = JoinLayers(integrated[["RNA"]])
```


## UMAP

```{r}
integrated  = RunUMAP(integrated , dims = 1:30, 
                      reduction = "integrated.rpca",  
                      reduction.name = "rpca-umap", 
                      reduction.key  = "RPCAUMAP_")
# Cluster
integrated = FindNeighbors(
  integrated,
  reduction  = "integrated.rpca",
  dims       = 1:30,
)
set.seed(2025)
integrated = FindClusters(integrated, resolution = 1)
integrated = FindClusters(integrated, resolution = 0.8)
integrated = FindClusters(integrated, resolution = 0.5)
integrated = FindClusters(integrated, resolution = 0.3)

saveRDS(integrated, file.path(output_dir,"integrated-rpca_BDF-ref.RDS"))

```

# Session information
```{r}
gc()

sessionInfo()
```

